Title here

### 2017-01-04 portfolio/text/ms.Rmd
More edits and references

Individuals relying on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, management, and
economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels fishing a greater diversity of species have less revenue
variability than those fishing a lower diversity. However, it is unclear if
these benefits extend to individual fishers and how year-to-year changes in
diversification affect revenue and revenue variability. Here, we identify three
axes by which fishers in Alaska, one of the world's most productive fishing
regions, can diversify fishing activities. We show that, despite increasing
specialization over the last 30 years, fishing permits with higher species
diversity reduces individual revenue variability and fishing an additional
permit is associated with higher revenue and lower variability. However,
increasing species diversity within the constraints of existing permits has
a fishery-dependent effect on revenue and is usually (`r g2_dens`% probability)
associated with increased revenue uncertainty the following year. Our results
demonstrate that the most effective option for individuals to decrease revenue
variability is to purchase permits to participate in additional or more diverse
fisheries. However, this option is expensive, often limited by regulations such
as catch share programs, and consequently unavailable to many individuals. With
increasing environmental variability in a changing climate, it will be
particularly important that individuals relying on natural resources for their
livelihood have effective strategies to reduce income variability.

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
findings within strategies, with respect to individual year-to-year behaviour,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, biological, or environmental
conditions change. For example, U.S. farmers specialized on wheat, corn, or
soybeans were hit harder than diversified farmers in the early 1980s when
a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource in a given year. For example,
downward economic shocks have been associated with increased diversity for
households relying on forests for income [@debela2012].

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, changes to fishing patterns will
sometimes be associated with a learning or adaptation phase as individuals
become accustomed to new locations, new gear, or new species [REF].
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers were forced to target other species or landed
more bycatch, adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our large-scale model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behaviour [@hilborn1985] and therefore
not included in our large-scale model, that might make a specific individual's
revenue less predictable if species diversity changes.

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial cost. Related barriers are apparent in other
natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries, while
catch shares [@costello2008] and limited season lengths [REF] can be effective
tools to avoid overfishing and distribute social capital, they can also limit
opportunities for diversification [@smith1986; @kasperski2013]. However, catch
shares, in particular, may also reduce income variability (Fig. S\ref{fig:ifq}A)
perhaps offsetting the effects of reduced diversification [@holland2016].

@@

TODO: finish this:
External market forces may interact in concert with management regulations to
promote specialization for fishers or other individuals dependent on natural
resources. With increasing environmental variability and extreme events in
a changing climate [@easterling2000], it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Though variability in individual revenue may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

>>>>

Individuals relying on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, management, and
economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels fishing a greater diversity of species have less revenue
variability than those fishing a lower diversity. However, it is unclear if
these benefits extend to the actions of individual fishers and how year-to-year
changes in diversification affect revenue and revenue variability. Here, we
identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities. We show that,
despite increasing specialization over the last 30 years, fishing permits with
higher species diversity reduces individual revenue variability and fishing an
additional permit is associated with higher revenue and lower variability.
However, increasing species diversity within the constraints of existing permits
has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
probability) associated with increased revenue uncertainty the following year.
Our results demonstrate that the most effective option for individuals to
decrease revenue variability is to purchase permits to participate in additional
or more diverse fisheries. However, this option is expensive, often limited by
regulations such as catch share programs, and consequently unavailable to many
individuals. With increasing environmental variability in a changing climate, it
will be particularly important that individuals relying on natural resources for
their livelihood have effective strategies to reduce income variability.

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
findings within strategies, with respect to individual year-to-year behaviour,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, local knowledge and experience are
a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
fishing patterns may sometimes be associated with a learning or adaptation phase
as individuals become accustomed to new locations, gear, or species.
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers were forced to target other species or landed
more bycatch, adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our large-scale model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
@branch2006] and therefore not included in our large-scale model, that might
make a specific individual's revenue less predictable if species diversity
changes.

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial expense. Related barriers are apparent in
other natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries,
regulations that you have been shown to reduce overfishing --- including
improved assessments [@melnychuk2017] and individual quota systems
[@costello2008] --- may ultimately restrict individual diversification
[@smith1986; @kasperski2013] or at least make developing a diverse fishing
portfolio more expensive. However, individual quota systems, in particular, may
also reduce income variability (Fig. S\ref{fig:ifq}A) perhaps offsetting the
effects of reduced diversification [@holland2016].

@@

TODO: finish this:
External market forces may interact in concert with management regulations to
promote specialization for fishers or other individuals dependent on natural
resources. With increasing environmental variability and extreme events in
a changing climate [@easterling2000], it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Although variability in individual revenue may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

<!--take home is the management context is key and controls show that
it's more complicated-->
